Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
fixed point notation

Let $\mathcal{F}\langle I,F \rangle(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.

The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}\langle I,F \rangle(x)$ be the same number represented in a fixed point domain as above. $\mathcal{F}\langle I,F \rangle(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands. 
$\mathcal{F}\langle I,F \rangle(c_1) \pm \mathcal{F}\langle I,F \rangle(c_2) = c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.

\item Multiplication: $\mathcal{F}\langle I,F \rangle(c_1) * \mathcal{F}\langle I,F \rangle(c_2) =  c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*\mathcal{F}\langle I,F \rangle(c_2)| + |\delta_{T2}*\mathcal{F}\langle I,F \rangle(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item In this implementation we only need to divide by positive numbers(where it is needed for checking stability). 
\end{enumerate}

\item {\bf Overflow:}
The maximum size of a real number $x$ that can be represented in a fixed point domain as $\mathcal{F}\langle I,F \rangle(x)$ is $\pm (2^{I-1}+1-2^{-F})$. Numbers outside this range cannot be represented by the domain. We assume no saturation (where any overflow is represented as the largest or smallest possible number), and check that overflow does not occur.

@@

We use $\mathcal{F}\langle I_c,F_c \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}\langle I_p,F_p \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model. Thus all mathematical operations occurring in the digital controller occur in the domain $\mathcal{F}\langle I_c,F_c \rangle$, and all other calculations are carried out in the domain $\mathcal{F}\langle I_p,F_p \rangle$.

We will use $\mathcal{F}\langle I_c,F_c \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}\langle I_p,F_p \rangle(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model ($I_p \geq I_c \wedge F_p \geq F_c$). Thus any mathematical operations occurring in the digital controller will occur in the range of $\mathcal{F}\langle I_c,F_c \rangle$, and all other calculations will be carried out in the range of $\mathcal{F}\langle I_p,F_p \rangle$.

>>>>

Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.

The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed point domain as above. $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands. 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.

\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item In this implementation we only need to divide by positive numbers(where it is needed for checking stability). 
\end{enumerate}

\item {\bf Overflow:}
The maximum size of a real number $x$ that can be represented in a fixed point domain as $\mathcal{F}_{\langle I,F \rangle}(x)$ is $\pm (2^{I-1}+1-2^{-F})$. Numbers outside this range cannot be represented by the domain. We assume no saturation (where any overflow is represented as the largest or smallest possible number), and check that overflow does not occur.

@@

We use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model. Thus all mathematical operations occurring in the digital controller occur in the domain $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations are carried out in the domain $\mathcal{F}_{\langle I_p,F_p \rangle}$.

We will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model ($I_p \geq I_c \wedge F_p \geq F_c$). Thus any mathematical operations occurring in the digital controller will occur in the range of $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations will be carried out in the range of $\mathcal{F}_{\langle I_p,F_p \rangle}$.
